Internet Attribute Certificate Profile for Authorization (RFC5755) - A Cryptographic Analyses

### 2013-05-29 paper.tex
Version 1.0

\begin{abstract}
ABSTRACT
\end{abstract}

@@

\section{Introduction}
\label{Introduction}
This paper outlines some of the cryptographic aspects offered by the Attribute Certificates\cite{rfc_ac} to the Public Key Infrastructure as defined by X.509\cite{rfc_x509}. Attribute Certificates (AC) is currently (may 2013) a \textit{Proposed Standard} with the IETF and is awaiting approval for the \textit{Draft Standard} category. The RFC5755 which covers AC's will obsolete an old IETF document describing AC's\cite{rfc_oldac}.

@@

 Figure~\ref{fig:diff} shows the differences between an AC and a PKC. As can be seen, ACs hold a value referring to the holder. The RFC~\cite{rfc_ac} prescripes three ways in which such an AC can be bound to a PKC, of which only one may be used per AC.
\begin{itemize}
	\item \textit{baseCertificateID}, the issuer and serial number of the holder's PKC. "For any environment where the AC is passed in an authenticated message or session and where the authentication is based on the use of an X.509 PKC, the Holder field SHOULD use the baseCertificateID."~\cite{rfc_ac}
	\item \textit{entityName}, the name of the claimant or role. This name must the same as on the PKC.
	\item \textit{objectDigestInfo}, used to directly authenticate the holder. This objectDigestInfo contains a hash of an object the AC is linked to.
\end{itemize}
Furthermore, it has a value containing the attributes belonging to this AC.

@@

\subsection{Examples of usage}
An AC can be used in different security related services. "As an example it could also be used to certify authorization information related to a public key. More specifically, it may be used to limit the liability resulting from a digital signature, or to constrain the usage of a public key (e.g., transaction of a limited value, certain time windows, etc.)."~\cite{tilborg2011encyclopedia}.

Different default attributes have been proposed, some of which are listed below.~\cite{benantar2006access,rfc_ac}
\begin{itemize}
	\item \textit{Service Authentication Information}, which identifies the AC holder to the service by name, with optional service-specific authentication information. This attribute could, by using some encryption scheme, also include the holder's identity and password.
	\item \textit{Charging identity}, which is used to charge the holder for the use of services.
	\item \textit{Role}, which is used to specify the role the holder has within a specific service.
	\item \textit{Clearance}, which defines the clearance the holder has for specific information (secret, top-secret, etc.).
	\item \textit{Group}, which describes the group the holder belongs to.
\end{itemize}

@@

\subsection{Public Key Infrastructure to fit AC in}
To use ACs within an organisation, there are mainly two implementation defaults, the so called \textit{push} and \textit{pull} models. This paper roughly describes the main arguments for choosing either of the two.

The \textit{push} model requires the AC holder to own his AC. When speaking to the Server, he/she is required to attach the AC with the request. "The 'push' model is suitable in application where the client's permissions should be authenticated/validated in the client's 'home' domain."~\cite{godavari2001secure} This means that no new connections need to be set up between the Server and other parties, as will be in the \textit{pull} model.

The \textit{pull} model requires the Server to retrieve the AC from a repository or an AC issuer. This means the implementation does not involve a change in the client-server protocol, which has benefit over the \textit{push} method when implemented within an existing infrastructure. "The 'pull' model is suitable when the client's priviliges should be authenticated in the inter-domain"~\cite{godavari2001secure} Furthermore, there is no need to update client's ACs when they expire, this can be done within the repository or AC issuer.

\section{Cryptography in Attribute Certificates}
\label{cryptography_in_ac}
This Section describes the role of cryptography within the use of ACs and the way ACs are constructed and issued by an AC issuer. Lets stress again that this section is not ment to give a complete overview of the cryptography involved with ACs.

@@

After the AC has been build, either the \textit{push} or \textit{pull} model can be used to either distribute the AC to the client, or to distribute it to the AC repository. AC verifiers will verify the validity of an AC for an action or authenticate using the following steps.~\cite{rfc_ac}
\begin{enumerate}
	\item The AC holder's PKC must be present, and the entire certification path of that PKC must be verified.
	\item The AC signature must be cryptographically correct and the issuers PKC certification path must be verified.
	\item The AC issuer must be trusted directly (as mentioned above, the ACs Public Key should be trusted).
	\item The time for which the AC is being evaluated must be within the AC validity.
	\item The AC must follow the standard.
	\item The AC must not contain unsupported critical extensions.
	\item The AC should not have been revoked.
\end{enumerate}

\subsection{Cryptography}
ACs are signed by the AA using a Signature Algorithm which is specified within the AC. The algorithms allowed for the signature are listed in different RFCs~\cite{rfc_alg1,rfc_alg2,rfc_alg3,rfc_alg4}.

\section{Security Requirements}
\label{security_requirements}
What are the security requirements outlined in this RFC? What security problems does it attempt to solve? What (cryptographic) attacks does it defend against?
--> revocation

>>>>

\begin{abstract}
This paper touches upon some of the cryptographic and PKI elements concerned with Attribute Certificates. These attributes are used to extend the lifetime and functionality of regular Public Key Certificates in a Public Key Infrastructure. The Attribute Certificates are binded to a Public Key Certificate, have a shorter lifetime (quite temporarily) and add attributes from specific services to the Public Key Certificate. The attributes support all sorts of functionality, including authorization, access regulation, charging mechanisms and much more. This paper draws some of the characteristics and usage of Attribute Certificates in a Public Key Infrastructure, aswell as providing the reader with some security requirements when working with them.
\end{abstract}

@@

\section{Introduction}
\label{Introduction}
This paper outlines some of the cryptographic aspects offered by the Attribute Certificates\cite{rfc_ac} to the Public Key Infrastructure as defined by X.509\cite{rfc_x509}. Attribute Certificates (AC) is currently (may 2013) a \texttt{Proposed Standard} with the IETF and is awaiting approval for the \texttt{Draft Standard} category. The RFC5755 which covers AC's will obsolete an old IETF document describing AC's\cite{rfc_oldac}.

@@

 Figure~\ref{fig:diff} shows the differences between an AC and a PKC. As can be seen, ACs hold a value referring to the holder. The RFC~\cite{rfc_ac} prescripes three ways in which such an AC can be bound to a PKC, of which only one may be used per AC.
\begin{itemize}
	\item \texttt{baseCertificateID}, the issuer and serial number of the holder's PKC. \textit{"For any environment where the AC is passed in an authenticated message or session and where the authentication is based on the use of an X.509 PKC, the Holder field SHOULD use the baseCertificateID."}~\cite{rfc_ac}
	\item \texttt{entityName}, the name of the claimant or role. This name must the same as on the PKC.
	\item \texttt{objectDigestInfo}, used to directly authenticate the holder. This objectDigestInfo contains a hash of an object the AC is linked to.
\end{itemize}
Furthermore, it has a value containing the attributes belonging to this AC.

@@

\subsection{Examples of usage}
An AC can be used in different security related services. \textit{"As an example it could also be used to certify authorization information related to a public key. More specifically, it may be used to limit the liability resulting from a digital signature, or to constrain the usage of a public key (e.g., transaction of a limited value, certain time windows, etc.)."}~\cite{tilborg2011encyclopedia}.

Different default attributes have been proposed, some of which are listed below.~\cite{benantar2006access,rfc_ac}
\begin{itemize}
	\item \texttt{Service Authentication Information}, which identifies the AC holder to the service by name, with optional service-specific authentication information. This attribute could, by using some encryption scheme, also include the holder's identity and password.
	\item \texttt{Charging identity}, which is used to charge the holder for the use of services.
	\item \texttt{Role}, which is used to specify the role the holder has within a specific service.
	\item \texttt{Clearance}, which defines the clearance the holder has for specific information (secret, top-secret, etc.).
	\item \texttt{Group}, which describes the group the holder belongs to.
\end{itemize}

@@

\subsection{Public Key Infrastructure to fit AC in}
To use ACs within an organisation, there are mainly two implementation defaults, the so called \texttt{push} and \texttt{pull} models. This paper roughly describes the main arguments for choosing either of the two.

The \texttt{push} model requires the AC holder to own his AC. When speaking to the Server, he/she is required to attach the AC with the request. \textit{"The 'push' model is suitable in application where the client's permissions should be authenticated/validated in the client's 'home' domain."}~\cite{godavari2001secure} This means that no new connections need to be set up between the Server and other parties, as will be in the \texttt{pull} model.

The \texttt{pull} model requires the Server to retrieve the AC from a repository or an AC issuer. This means the implementation does not involve a change in the client-server protocol, which has benefit over the \texttt{push} method when implemented within an existing infrastructure. \textit{"The 'pull' model is suitable when the client's priviliges should be authenticated in the inter-domain"}~\cite{godavari2001secure} Furthermore, there is no need to update client's ACs when they expire, this can be done within the repository or AC issuer.

Figure~\ref{fig:PKI_models} shows the different protocol possibilities within one figure. It shows the different possible actors (AC issuer, Client or AC Holder, Server or AC verifier and Repository) and their possible communication flows.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{PKI_models.png}
	\caption{Protocol communication for \textit{push} and \textit{pull} models~\cite{rfc_ac}}
	\label{fig:PKI_models}
\end{figure}

\section{Cryptography in Attribute Certificates}
\label{cryptography_in_ac}
This Section describes the role of cryptography within the use of ACs and the way ACs are constructed and issued by an AC issuer. Lets stress again that this paper is not ment to give a complete overview of the cryptography involved with ACs and the author is in no way responsible for any damage caused by the information in this paper.

@@

After the AC has been build, either the \texttt{push} or \texttt{pull} model can be used to either distribute the AC to the client, or to distribute it to the AC repository. AC verifiers will verify the validity of an AC for an action or authenticate using the following steps~\cite{rfc_ac}.
\begin{enumerate}
	\item The AC holder's PKC must be present, and the entire certification path of that PKC must be verified.
	\item The AC signature must be cryptographically correct and the issuers PKC certification path must be verified.
	\item The AC issuer must be trusted directly (as mentioned above, the ACs Public Key should be trusted).
	\item The time for which the AC is being evaluated must be within the AC validity.
	\item The AC must follow the standard.
	\item The AC must not contain unsupported critical extensions.
	\item The AC should not have been revoked. (See Section~\ref{security_requirements})
\end{enumerate}

\subsection{Cryptography}
ACs are signed by the AA using a Signature Algorithm which is specified within the AC. The algorithms allowed for the signature are listed in different RFCs~\cite{rfc_alg1,rfc_alg2,rfc_alg3,rfc_alg4}. The RFC~\cite{rfc_ac} states \textit{"The binding between an AC holder and attributes cannot be stronger than the cryptographic module implementation and algorithms used to generate the signature.  Short key lengths or weak hash algorithms will limit the utility of an AC. AC issuers are encouraged to note advances in cryptology so they can employ strong cryptographic techniques."} Which states clear that the responsibility for using cryptographic hash functions lies with the AA.

The other cryptography that could be involved in ACs is the placement of login details (username and password) of an AC owner in the AC itself. In this case, ofcourse measures have to be taken to avoid plaintext passwords in the AC. The RFC~\cite{rfc_ac} let you decide upon proper cryptographic encryption functions necessary for the protection of the login credentials. The RFC~\cite{rfc_ac} suggests the use of Public Key Cryptography (encrypting with the Public Key of the Server or AC verifier). It is again the case that the AA is responsible for good cryptographic management of the issues mentioned previously.

\section{Security Requirements}
\label{security_requirements}
This section discusses the requirements set for the AC (which AC all meets) and discusses the security requirements necessary for securely operating with ACs. Furthermore it will touch upon AC revocation.

\subsection{AC Requirements}
The RFC~\cite{rfc_ac} enumerates a set of requirements they meet with this RFC.
\begin{enumerate}
	\item \textit{"Support for short-lived as well as long-lived ACs. Typical short-lived validity periods might be measured in hours, as opposed to months for PKCs. Short validity periods allow ACs to be useful without a revocation mechanism."} This is met by forceing the use of a validity value on an AC which can be set \texttt{YYYYMMDDHHMMSSZ}.
	\item \textit{"Issuers of ACs should be able to define their own attribute types for use within closed domains."} This is met by allowing attributes to be made by the AAs themselves, though, keeping consistency by forceing the use of some predefined way of defining them.
	\item \textit{"Some standard attribute types, which can be contained within ACs, should be defined. Examples include 'access identity', 'group', 'role', 'clearance', 'audit entity', and 'charging identity'"} This is met by having these implemented as defaults.
	\item \textit{"Standard attribute types should be defined in a manner that permits an AC verifier to distinguish between uses of the same attribute in different domains. For example, the 'Administrators group' as defined by 'Baltimore' and the 'Administrator group' as defined by 'SPYRUS' should be easily distinguished."} This is met by allowing an optional \texttt{roleAuthority} for every \texttt{roleName}.
	\item \textit{"It should be possible to 'target' an AC at one, or a small number of, servers. This means that a trustworthy non-target server will reject the AC for authorization decisions."} This is met by the implementation of so called \texttt{Target} and \texttt{TargetCert} fields.
	\item \textit{"ACs should be defined so that they can either be 'pushed' by the client to the server, or 'pulled' by the server from a repository or other network service, including an online AC issuer."} This is met by the two possible implementation defaults for the Server or AC verifier.
\end{enumerate}

\subsection{Security Requirements for Safe Use}
Several security requirements for safe use of ACs are recommended by the RFC~\cite{rfc_ac}. This subsection discusses most of them. For a complete view of the requirements, the reader is reffered to the RFC~\cite{rfc_ac}.

\begin{itemize}
	\item Masquerading of the AA after the private keys have been compromised could lead to serious security problems. The ACs that are created by an attacker are checked offline a Server or AC verifier, and therefore can cause serious damage. Revocation of the AAs public key (trust in the AA) should be performed to overcome this problems. Detection could be quite hard, since offline verification is done. This is especially dangerious if the \texttt{push} model is used, in which a rogue client can easily send his rogue AC to the AC verifier.

	\item If an AA loses his private keys, this also leads to serious problems. All revocation lists have to be signed by the AA, and therefore, revocation becomes a critical problem.

	\item AAs are recommended to have an online \texttt{AC Revocation Status Service} by which an AC verifier can easily check the status of an AC. The freshness of the AC Status highly reduces vulnerabilites since these are detected quicker.

	\item Cryptographic hash functions, signing algorithms, public key cryptography, keylengths, etc. all are important since \textit{"The binding between an AC holder and attributes cannot be stronger than the cryptographic module implementation and algorithms used to generate the signature."}~\cite{rfc_ac}.

	\item The RFC~\cite{rfc_ac} is really clear about how and what to check for the binding between the AC and the PKC. An AC verifier should strictly follow the protocol described by the RFC. Not following protocol, especially concering the fields for \texttt{entityName} and \texttt{baseCertificateID} could lead to serious security issues.

	\item It is highly recommended to not use AC chains and to make the signing AA for an AC directly trustworthy.
\end{itemize}

\subsection{AC Revocation}
This paper mentioned AC revocation several times, this Section will explain the rough base for AC revocation as set by the RFC~\cite{rfc_ac}. Typically, the use of ACs is ment \textit{temporarily}, and therefore revocation should occur less often than one sees within a PKC-like environment. Though, since situations (like some mentioned in the previous subsection) could occur in which revocation is desired, the AC standard allows for mechanisms to revoke ACs. Two mechanisms exist.

The \texttt{Never revoke} scheme MUST be supported by all AC verifiers. This scheme lets ACs contain a possible \texttt{noRevAvail} field, which implicitly means that the status of the AC can be verified at the AA. This particular AC with the \texttt{noRevAvail} field set to true, will never end up on any revocation list and should therefore be checked at the AA.

The \texttt{Pointer in AC} scheme allows for a \textit{pointer} to a revocation list or a source of revocation status information in the AC. The ACs that have fields set to use this scheme can end up at those lists.

It is forbidden for an AC to be in both schemes. If an AC verifier only supports the \texttt{Never revoke} scheme, then all ACs who are using the \texttt{Pointer in AC} scheme must be rejected, since there status cannot be checked online (and the AC verifier does not support the revocation list pointers).
